---
title: "02_variable-selection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("install.load")) {
  install.packages("install.load")
  library(install.load)
}

# Packages
list.of.packages <- c("tidyverse", "lme4", "lmtest", "nlme", "lcmm", "MASS",
                      "brms", "sjPlot", "sjstats")

install.load::install_load(list.of.packages)
```

```{r}
load("../data/processed/model-df.Rdata")

x <- model.trans
```


```{r}
ctrl <- lmeControl(opt = 'optim')
lme00 <- lme(Akk.logit ~ il17.logit,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme00)
```

This shows IL17 to be a significant predictor of Akk.logit levels.

However, the model seriously deviates from assumptions of normality; there is notable patterning to the residuals.

```{r}
plot(lme00)
```

Regardless, I'll use this framework to estimate predictor variables to find the best performing lme model and then test its prediction accuracy against other model types. 

The next model I'll check includes infection as a covariate. 

```{r}
lme01 <- lme(Akk.logit ~ il17.logit + Infection,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme01)
```

Both il17.logit and Infection are significant, as is Infection's effect on the overall model by likelihood ratio test.

```{r}
lrtest(lme00, lme01)
```

Now I'll check if there is an interaction between the covariates. 

```{r}
lme02 <- lme(Akk.logit ~ il17.logit * Infection,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme02)
lrtest(lme01, lme02)
```

Interaction term is not significant, nor does the inclusion of an interaction term reduce the residual variance of the model. 


Now I'll include `Season`, which is a factor with four variables.

```{r}
lme03 <- lme(Akk.logit ~ il17.logit + Infection + Season,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme03)
```

None of the seasons are significant in this model.

```{r}
lme04 <- lme(Akk.logit ~ il17.logit + Infection + days,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme04)
```

The days are highly signficant, and now il17.logit has a pvalue of 0.09. Given that we'd like to include it in the model, the alpha threshold for variable selection will be defined as 0.1.

```{r}
lrtest(lme04, lme01)
```

The inclusion of days significantly improves the model.

```{r}
lme05 <- lme(Akk.logit ~ il17.logit * days + Infection,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme05)
```

No significant interaction between il17.logit and time -- exclude from the model.

```{r}
lme06 <- lme(Akk.logit ~ il17.logit + Infection + days + days2,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme06)
lrtest(lme06, lme04)
```

A second-order time component benefits the model, though with a pvalue > 0.05. In order to include il17.logit the alpha must be set to 0.1; applying that threshold to this effect includes days2.

```{r}
lme07 <- lme(Akk.logit ~ il17.logit + Infection + days + days2 + days3,
             random = list(~1 | SubjectID),
             data = x, na.action = na.omit,
             method = "ML", control = ctrl)
summary(lme07)
```

A cubic time term does not improve the model.

Now I'll check for a random slope term for days.

```{r}
lme04.1 <- lme(Akk.logit ~ il17.logit + Infection + days,
               random = list(~1 + days | SubjectID),
               data = x, na.action = na.omit,
               method = "ML", control = ctrl)
summary(lme04.1)
lrtest(lme04, lme04.1)
```

Adding in a random slope as well as intercept significantly improves the model. 

```{r}
lme04.1 <- lme(Akkermansia ~ il17.logit + Infection + days,
               random = list(~1 +days | SubjectID),
               data = x, na.action = na.omit,
               method = "ML", control = ctrl)
summary(lme04.1)
```{r}
ctrl <- lmeControl(opt = 'optim')

lm00 <- lme(Akk.logit ~ il17.logit + Infection + Season + days,
            random = list(~1 | SubjectID),
            data = df, na.action = na.omit,
            method = "ML", control = ctrl)
summary(lm00)

lm00 <- lmer(Akk.logit ~ il17.logit + Infection + Season + days + (1 | SubjectID),
             data = df, na.action = na.omit)
summary(lm00)

```

```{r}
glm00 <- glmer.nb(Akkermansia ~ IL17F + Infection + days + (1|SubjectID),
                  data = x)

summary(glm00)
```


```{r}
brm00 <- brms::brm(Akkermansia ~ IL17F + Infection + days + (1|SubjectID),
                  data = x, sparse = TRUE, family = negbinomial())
summary(brm00)

plot_model(brm00)
plot(brm00)
plot(marginal_effects(brm00))

```

```{r}
brm01 <- brms::brm(Akkermansia ~ IL17F + Infection + (1|SubjectID),
                  data = x, sparse = TRUE, family = negbinomial())
summary(brm01)

```

```{r}
brm02 <- brms::brm(Akkermansia ~ IL17F + Infection + Season + (1|SubjectID),
                  data = x, sparse = TRUE, family = negbinomial())
summary(brm02)

```
```{r}
brm03 <- brms::brm(Akkermansia ~ IL17F + Infection + cluster + (1|SubjectID),
                  data = x, sparse = TRUE, family = negbinomial())
summary(brm03)
```


